---
title: "Introduction to acor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to acor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(acor)
```

## Overview

`acor` provides asymmetric correlation measures for settings where `Y` is the
outcome and `X` is a predictor. The package also includes several familiar
symmetric correlation measures so that directional and non-directional analyses
can be compared in one interface.

The main entry points are:

- `acor()` to compute correlation estimates
- `acor.test()` to perform asymptotic inference

## Available measures

The asymmetric measures are:

- `akc`: Asymmetric Kendall Correlation
- `cid`: Concordance-Discordance Index, equal to `(AKC + 1) / 2`
- `agc`: Asymmetric Grade Correlation
- `cma`: Coefficient of Monotone Association, equal to `(AGC + 1) / 2`

For `akc` and `agc`, independence corresponds to `0`. For `cid` and `cma`,
independence corresponds to `0.5`.

The package also supports symmetric measures:

- `tau_a`, `tau_b`
- `gamma`
- `rho_a`, `rho_b`
- `pearson`

## Basic usage

```{r}
set.seed(1)
x <- rnorm(100)
y <- 0.6 * x + rnorm(100)

acor(x, y, method = "akc")
acor(x, y, method = "cid")
acor(x, y, method = "pearson")
```

## Hypothesis tests

`acor.test()` supports all measures listed above. Default variance is
`variance = "ij"` (infinitesimal jackknife) for `akc`, `agc`, `cid`, `cma`,
`tau_b`, `gamma`, and `rho_b`; use `variance = "plugin"` for closed-form
kernels (required for `pearson`, `tau_a`, and `rho_a`, or as an alternative
where IJ is available).

```{r}
set.seed(2)
x <- rnorm(120)
y <- 0.5 * x + rnorm(120)

acor.test(x, y, method = "akc")
acor.test(x, y, method = "tau_b")
acor.test(x, y, method = "pearson", variance = "plugin")
acor.test(x, y, method = "cma", alternative = "greater")
```

By default, inference is performed under IID assumptions. For
time series, set `IID = FALSE` to use a HAC variance estimator.

```{r}
set.seed(3)
n <- 120
x <- cumsum(rnorm(n))
y <- 0.3 * x + cumsum(rnorm(n))

acor.test(x, y, method = "agc", IID = FALSE)
```

## Multiple predictors

If `X` is a matrix, `acor()` returns one estimate per column. For
`acor.test()`, the null hypothesis is that all predictor-specific correlations
with `Y` are equal.

```{r}
set.seed(4)
y <- rnorm(150)
X <- cbind(
  x1 = y + rnorm(150, sd = 0.4),
  x2 = 0.5 * y + rnorm(150, sd = 0.8),
  x3 = rnorm(150)
)

acor(X, y, method = "agc")

fit <- acor.test(X, y, method = "akc")
fit
fit$results
fit$pairwise_results
```

## Interpretation

The asymmetric measures are directional: exchanging `X` and `Y` generally
changes the result. This is most useful when one wants to quantify the
association of one or more predictors with a designated outcome.

The rescaled measures `cid` and `cma` lie on `[0, 1]`, whereas `akc` and `agc` lie on
`[-1, 1]`.
